---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# statgenQTLxT

**statgenQTLxT** is an R package for fast multi-trait and multi-trial Genome Wide Association Studies (GWAS).

statgenQTLxT has extensive options for summarizing and visualizing results. It builds on the `statgenGWAS` package (for single trait GWAS) which is available from [CRAN](https://biometris.github.io/statgenGWAS/). The package uses data structures and plots defined in the `statgenGWAS` package. It is recommended to read the vignette of this package, accessible in R via `vignette(package = "statgenGWAS")` or online at <https://biometris.github.io/statgenGWAS/articles/GWAS.html> to get a general idea of those. 

## Installation

* Install from CRAN:

```{r, eval = FALSE}
install.packages("statgenQTLxT")
```

* Install latest development version from GitHub (requires [remotes](https://github.com/r-lib/remotes) package):

```{r, eval = FALSE}
remotes::install_github("Biometris/statgenQTLxT", ref = "develop", dependencies = TRUE)
```

