---
title: "Introduction to the statgenPipeline package"
author: "Bart-Jan van Rossum"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: false
    number_sections: true
bibliography: bibliography.bib    
vignette: >
  %\VignetteIndexEntry{Introduction to the statgenPipeline package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 4)
)
library(statgenPipeline)
options(width = 100, digits = 2)
```

The statgenPipeline package has been designed for performing multi-trait Genome Wide Association Studies (GWAS). It builds on the statgenGWAS package (for single trait GWAS) which is available from CRAN (https://cran.r-project.org/web/packages/statgenGWAS/index.html). The package uses data structures and plots defined in the stagenGWAS package. It is recommended to read the vignette of this package, accessible in R via `vignette(package = "statgenGWAS")` to get a general idea of those.

----

# The class `gData`

Data for analysis on genomic data comes from different sources and is stored in one data object of class **gData** (**g**enomic **Data**) for convenience. A gData object will contain all data needed for performing analyses, so the first thing to do when using the statgenPipeline package is creating a gData object. A gData object consists of different types of data. All data is optional, however for certain analyses, certain data is required. Additional data can always be added to an existing object later on.  
A gData object can be created using the function **_createGData_** from the statgenGWAS package. If an existing gData object is provided as an input to this function, data is added to the existing object and existing data will be overwritten with a warning if necessary. All data is checked for consistency and, if needed, default names are added with a warning message.  
A gData object created this way will consist of the following components: 

* Marker map, a data.frame describing the physical positions of the markers on the chromosomes,
* Marker matrix, a numerical matrix containing the genotyping,
* Phenotypic data, can either be a single data.frame or a list of data.frames,
* Kinship matrix,  describing the genetic relatedness between the different genotypes, 
* Further covariates that can be used in the analyses.

For more details see the statgenGWAS package.

----

# Multi-Trait GWAS

## Theoretical background

*Not sure we need more than the application here.*

## Application in the package

Multi-trait GWAS in the statgenPipeline package estimates the effect of a SNP in different trials or on different traits, one SNP at a time. Genetic and residual covariances are fitted only once, for a model without SNPs. Following the diagonalization scheme of [@Zhou2014], the following model is fit

$Y = \left(\begin{array}{c} Y_1 \\ \vdots \\ Y_p\end{array}\right) = \left(\begin{array}{c} X_1\gamma_1 \\ \vdots \\ X_p\gamma_p\end{array}\right) + \left(\begin{array}{c} x_1\beta_1 \\ \vdots \\ x_p\beta_p\end{array}\right) + \left(\begin{array}{c} G_1 \\ \vdots \\ G_p\end{array}\right) + \left(\begin{array}{c} E_1 \\ \vdots \\ E_p\end{array}\right)$

where $Y$ is a $np × 1$ vector of phenotypic values for $n$ genotypes and $p$ traits or trials. $x$ is the $n × 1$ vector of scores for the marker under consideration, and $X$ the $n × q$ design matrix for the other covariates. By default only a trait (environment) specific intercept is included. The vector of genetic background effects $\left(\begin{array}{c}G_1 \\ \vdots \\ G_p\end{array}\right)$ is Gaussian with zero mean and covariance $V_g \otimes K$, where $V_g$ is a $p × p$ matrix of genetic (co)variances, and $K$ an $n × n$ kinship matrix. Similarly, the residual errors ($\left(\begin{array}{c}E_1 \\ \vdots \\ E_p\end{array}\right)$) have covariance $V_e \otimes I_n$, for a $p × p$ matrix $V_e$ of residual (co)variances.

### Hypotheses for the SNP-effects {#common}

For each SNP, the null-hypothesis $\beta_1 = \dots = \beta_p = 0$ is tested, using the likelihood ratio test (LRT) described in [@Zhou2014]. If `estCom = TRUE`, additional tests for a common effect and for QTL-by-trait (QTL × T) or QTL-by-environment (QTL × E) are performed, using the parameterization $\beta_j = \alpha + \alpha_j (1 \leq j \leq p)$. As in [@Korte2012], we use likelihood ratio tests, but not restricted to the bivariate case. For the common effect, we fit the reduced model \eqn{\beta_j = \alpha}, and test if $\alpha = 0$. For the interactions, we test if $\alpha_1 = \dots = \alpha_p = 0$.

### Models for the genetic and residual covariance {#vc}

$V_g$ and $V_e$ can be provided by the user (`fitVarComp = FALSE`); otherwise one of the following models is used, depending on `covModel`. If `covModel = "unst"`, an unstructured model is assumed, as in [@Zhou2014]: $V_g$ and $V_e$ can be any positive-definite matrix, requiring a total of $p(p+1)/2$ parameters per matrix. If `covModel = "fa"`, a factor-analytic model is fitted using an
EM-algorithm, as in [@Millet2016]. $V_g$ and $V_e$ are assumed to be of the form $W W^t + D$, where $W$ is a $p × m$ matrix of factor loadings and $D$ a diagonal matrix with trait or environment specific values. $m$ is the order of the model, and the parameters $mG$ and $mE$ specify the order used for respectively $V_g$ and $V_e$. `maxIter` sets the maximum number of iterations used in the EM-algorithm. Finally, if `covModel = "pw"`, $V_g$ and $V_e$ are estimated 'pairwise', as in [@Furlotte2015]. Looping over pairs of traits or trials $1 \leq j < k \leq p$, $V_g[j,k] = V_g[k,j]$ and $V_e[j,k] = V_e[k,j]$ are estimated assuming a bivariate mixed model. The diagonals of $V_g$ and $V_e$ are fitted assuming univariate mixed models. If the resulting $V_g$ or $V_e$ is not positive-definite, they are replaced by the nearest positive-definite matrix. In case `covModel = "unst"` or `"pw"` it is possible to assume that $V_e$ is diagonal (`VeDiag = TRUE`).

----

# Worked example: multiple traits in one trial

As an example of the functionality of the package a worked example is provided using maize data from the European Union project DROPS. The data is available from https://data.inra.fr/dataset.xhtml?persistentId=doi:10.15454/IASSTN [@Millet2019] and the relevant data sets are included as data.frames in the statgenGWAS package.

## **Load data**

The first step is loading the data into R. 
```{r loadData} 
library(statgenGWAS)
data(dropsMarkers)
data(dropsMap)
data(dropsPheno)
```
dropsMarkers contains the coded marker information for 41,722 SNPs and 246 genotypes. dropsMap contains information about the positions of those SNPs on the B73 reference genome V2. dropsPheno contains data for the genotypic means (Best Linear Unbiased Estimators, BLUEs), with one value per experiment per genotype, for eight traits. For a more detailed description of the contents of the data see ```help(createGData, package = statgenGWAS)```.

## **Create gData object**

The next step is to create a **gData** object from the raw data that can be used for the GWAS analysis. For this the raw data has to be converted to a suitable format for a **gData** object, see ```help(dropsData, package = statgenGWAS)``` and the statgenGWAS vignette for more details.

```{r convertMarkers, echo=FALSE}
## Add genotypes as row names of dropsMarkers and drop Ind column.
rownames(dropsMarkers) <- dropsMarkers[["Ind"]]
dropsMarkers <- dropsMarkers[colnames(dropsMarkers) != "Ind"]
```

```{r convertMap, echo=FALSE}
## Add genotypes as row names of dropsMap.
rownames(dropsMap) <- dropsMap[["SNP.names"]]
## Rename Chomosome and Position columns.
colnames(dropsMap)[match(c("Chromosome", "Position"), colnames(dropsMap))] <- c("chr", "pos")
```

When running a multi-trait/multi-environment GWAS, all traits used in the analysis should be in the same data.frame. Hence, for using multiple traits in one environment, the **gData** object should contain at least one data.frame with several traits measured in one environment with the following format:

| genotype | Trait1 | Trait2 | Trait3 |
|:---------|:-------|:-------|:-------|
| G1 | 0.3 | 17 | 277 |
| G2 | 0.4 | 19 | 408 |
| G3 | 0.5 | 17 | 206 |
| G4 | 0.7 | 13 | 359 |

For using one trait in multiple trials (environments), the **gData** object should contain at least one data.frame with one trait measured in several environments with the following format:

| genotype | Trait1-Trial1 | Trait1-Trial2 | Trait1-Trial3 |
|:---------|:-------|:-------|:-------|
| G1 | 0.3 | 0.7 | 0.5 |
| G2 | 0.4 | 0.9 | 0.1 |
| G3 | 0.5 | 0.8 | 0.2 |
| G4 | 0.7 | 0.5 | 0.4 |

In this example, we will use several traits measured in one environment. ´dropsPheno´ contains genotypic means for 28 trials. To be able to run a GWAS analysis per trial the data has to be added as a list of 28 data.frames. These data.frames should have "genotype" as their first column and may only contain traits after that. Other columns need to be dropped.    

```{r addPheno}
## Convert phenotypic data to a list.
dropsPhenoList <- split(x = dropsPheno, f = dropsPheno[["Experiment"]])
## Rename Variety_ID to genotype and select relevant columns.
dropsPhenoList <- lapply(X = dropsPhenoList, FUN = function(trial) {
  colnames(trial)[colnames(trial) == "Variety_ID"] <- "genotype"
  trial <- trial[c("genotype", "grain.yield", "grain.number", "anthesis", 
                   "silking", "plant.height", "ear.height")]
  return(trial)
})
```
Note that if the phenotypic data consists of only one trial/experiment, it can be added as a single data.frame without first converting it to a list. In this case createGData will convert the input to a list with one item.

Now a gData object containing map marker information and phenotypes can be created. Kinship matrix and covariates may be added later on.
```{r createGdata}
## Create a gData object containing map, marker and phenotypic information.
gDataDrops <- createGData(geno = dropsMarkers,
                          map = dropsMap, 
                          pheno = dropsPhenoList)
```

To get an idea of the contents of the data a summary of the gData object can be made. This will give an overview of the content of the map and markers and also print a summary per trait per trial. Since there are 28 trials and 8 traits in gDataDrops giving a very large summary, the output is restricted here to one trial using the `trials` parameter of the summary function.
```{r sumGData}
## Summarize gDataDrops.
summary(gDataDrops, trials = "Mur13W")
```

## **Recoding and cleaning of markers**

Marker data has to be numerical and without missing values in order to do GWAS analysis. This can be achieved using the *codeMarkers* function. The marker data available for the DROPS project has already been converted from A/T/C/G to 0/1/2. The *codeMarkers* function in the package can still be used to further clean the markers, e.g. by removing the duplicate SNPs.
```{r removeDupMarkers}
## Remove duplicate SNPs from gDataDrops.
gDataDropsDedup <- codeMarkers(gDataDrops, impute = FALSE, verbose = TRUE) 
```

Note that duplicate SNPs are removed at random. To get reproducible results make sure to set a seed.

To demonstrate the options of the *codeMarkers* function, see ```help(codeMarkers, package = statgenGWAS)``` and the statgenGWAS vignette for more details.


## **Multi-trait GWAS**

The cleaned gData object can be used for performing multi-trait GWAS analysis. In this example the trial *Mur13W* is used to demonstrate the options of the *runMultiTraitGwas* function. 

*Add here the option with trait selection*
```{r mtg}
## Run multi-trait GWAS for xx traits in trial Mur13W.
GWASDrops <- runMultiTraitGwas(gData = gDataDropsDedup, 
                               trials = "Mur13W", 
                               covModel = "fa")
```

The output of the *runMultiTraitGwas* function is an object of class GWAS. This is a list consisting of five components described below.  
**GWAResult**: a list of data.tables, one for each trial for which the analysis was run. Each data.table has the following columns:

| | |
|:-----|:--------------------------------------------------|
| snp | SNP name |
| trait | trait name |
| chr | chromosome on which the SNP is located |
| pos | position of the SNP on the chromosome |
| pValue | P-value for the SNP |
| LOD | LOD score for the SNP, defined as $-\log_{10}(pValue)$ |
| effect | effect of the SNP on the trait value |
| effectSe | standard error of the effect of the SNP on the trait value |
| allFreq | allele frequency of the SNP |
```{r gwaRes}
head(GWASDrops$GWAResult$Mur13W)
```

**signSnp**: a list of data.tables, one for each trial for which the analysis was run, containing the significant SNPs. Optionally also the SNPs close to the significant SNPs are included in the data.table. See [**3.6.7**](#sign) for more information on how to do this. The data.tables in signSnp consist of the same columns as those in GWAResult described above. Two extra columns are added:

| | |
|:-----|:--------------------------------------------------|
| snpStatus | either "significant SNP" or "within ... of a significant SNP" |
| propSnpVar | proportion of the variance explained by the SNP |
```{r signSnp}
print(GWASDrops$signSnp$Mur13W, row.names = FALSE)
```

**kinship**: the kinship matrix (or matrices) used in the GWAS analysis. This can either be the user provided kinship matrix or the kinship matrix computed when running the *runMultiTraitGwas* function.

**thr**: a list of thresholds, one for each trial for which the analysis was run, used for determining significant SNPs. 

**GWASInfo**: additional information on the analysis, e.g. the call and the type of threshold used.

### GWAS Summary

For a quick overview of the results, e.g. the number of significant SNPs, use the summary function.
```{r sumMtg}
## Create summary of GWASDrops.
summary(GWASDrops)
```

### GWAS Plots

The plot.GWAS function can be used to visualize the results in GWASDrops, with a QQ-plot, Manhattan plot or QTL-plot. More details for each `plotType` are available in the statgenGWAS vignette.


#### QQ plots

A QQ-plot of the observed against the expected $-\log_{10}(p)$ values can be made by setting `plotType = "qq"`. Most of the SNPs are expected to have no effect, resulting in P-values uniformly distributed on $[0,1]$, and leading to the identity function ($y=x$) on the $-\log_{10}(p)$ scale. As in the plot below, deviations from this line should only occur on the right side of the plot, for a small number of SNPs with an effect on the phenotype (and possibly SNPs in LD). There is **inflation** if the observed $-\log_{10}(p)$ values are always above the line $y=x$, and (less common) **deflation** if they are always below this line. A QQ-plot thefore gives a first impression of the quality of the GWAS model: if for example $-\log_{10}(p)$ values are consistently too large (inflation), the correction for genetic relatedness may not be adequate.
```{r qqMtg} 
## Plot a qq plot of GWAS Drops.
plot(GWASDrops, plotType = "qq")
```

#### Manhattan plots

A manhattan plot is made by setting `plotType = "manhattan"`. Significant SNPs are marked in red.
```{r manhattanMtg}
## Plot a manhattan plot of GWAS Drops.
plot(GWASDrops, plotType = "manhattan")
```

More options linked with `plotType = "manhattan"` are described in the statgenGWAS vignette.

#### QTL plots

A qtl plot can be made by setting `plotType = "qtl"`. In this plot the significant SNPs are marked by circles at their genomic positions, with diameter proportional to the estimated effect size; for an example see @Millet2016. Typically, this is done for multiple traits or environments, with the genomic position on the x-axis, which are displayed horizontally above each other and can thus be compared.    

Since the traits are measured on a different scale, the effect estimates cannot be compared directly. For better comparison, one can set `normalize = TRUE`, which divides the estimates by the standard deviation of the phenotype.

```{r qtlMtgNorm}
## Plot a qtl plot of GWAS Drops for Mur13W.
## Set significance threshold to 5 and normalize effect estimates.
plot(GWASDrops, plotType = "qtl", yThr = 5, normalize = TRUE)
```

Other parameters can be used to plot a subset of the chromosomes (`chr`) and directly export the plot to .pptx (`exportPptx = TRUE` and specify `pptxName`). Note that the officer package is required for this. A full list of parameters can be found by running `help(plot.GWAS)`. 

## Kinship matrices

The *runMultiTraitGwas* function has an argument `kinshipMethod`, which defines the kinship matrix used for association mapping. Kinship matrices can be computed directly using the *kinship* function or within the *runMultiTraitGwas* function. There are four options: (1) using the covariance between the scaled SNP-scores (`kinshipMethod = "astle"`, the default; see e.g. equation (2.2) in @Astle2009) (2) Identity by State (`kinshipMethod = "IBS"`; see e.g. equation (2.3) in @Astle2009) (3) using the formula by @VanRaden2008 (`kinshipMethod = "vanRaden"`) (4) User-defined, in which case the parameter `kin` needs to be specified.

By default, the same kinship matrix is used for testing all SNPs (`GLSMethod = "single"`). When `GLSMethod = "multi"`, the kinship matrix is chromosome-specific. In this case, the function fits variance components and computes effect-estimates and P-values for each chromosome in turn, using the kinship matrix for that chromosome. Each chromosome-specific kinship matrix is computed using the method specified by the argument `kinshipMethod`, using all SNPs that are **not** on this chromosome. As shown by @Rincent2014, this often gives a considerable improvement in power.

```{r mtgChrSpec, eval = FALSE}
## Run multi-trait GWAS for trial 'Mur13W'.
## Use chromosome specific kinship matrices computed using method of van Raden.
GWASDropsChrSpec <- runMultiTraitGwas(gData = gDataDropsDedup, 
                                      trials = "Mur13W",
                                      GLSMethod = "multi",
                                      kinshipMethod = "vanRaden",
                                      covModel = "fa")
```

# Worked example: one trait measured in multiple trials

## **Create gData object**

In this example, we will focus on one trait, grain yield, in all trials. ´dropsPheno´ contains genotypic means for 28 trials. To be able to run a GWAS analysis with one trait measured in all trials, the data has to be reshaped and added as a single data.frame with "genotype" as first column and traits after that.    

```{r addPhenoxE}
## Reshape phenotypic data to data.frame in wide format containing only grain yield.
PhenoDat <- reshape(dropsPheno[,c("Experiment","Variety_ID","grain.yield")], 
                    timevar = "Experiment", 
                    idvar = "Variety_ID", 
                    direction = "wide", 
                    v.names = "grain.yield")
## Rename Variety_ID to genotype and other columns with the trial name only.
names(PhenoDat)[1] <- "genotype"
names(PhenoDat)[2:ncol(PhenoDat)] <-  gsub("grain.yield.","",
                                           names(PhenoDat)[2:ncol(PhenoDat)] )
```

Now a gData object containing map marker information and phenotypes can be created. Kinship matrix and covariates may be added later on.
```{r createGdataxE}
## Create a gData object containing map, marker and phenotypic information.
gDataDropsxE <- createGData(geno = dropsMarkers,
                            map = dropsMap, 
                            pheno = PhenoDat)
summary(gDataDropsxE)
```

## **Recoding and cleaning of markers**

```{r removeDupMarkersxE}
## Remove duplicate SNPs from gDataDrops.
gDataDropsDedupxE <- codeMarkers(gDataDropsxE, impute = FALSE, verbose = TRUE) 
```

## **Multi-trait GWAS**

*Add here the option with trait selection*
```{r mtgxE}
## Run multi-trait GWAS for xx traits in trial Mur13W.
GWASDropsxE <- runMultiTraitGwas(gData = gDataDropsDedupxE, 
                                 covModel = "fa")
```


```{r gwaResxE}
head(GWASDropsxE$GWAResult$PhenoDat)
```

```{r signSnpxE}
head(GWASDropsxE$signSnp$PhenoDat, row.names = FALSE)
```


### GWAS Summary

For a quick overview of the results, e.g. the number of significant SNPs, use the summary function.
```{r sumMtgxE}
summary(GWASDropsxE)
```

### GWAS Plots

The plot.GWAS function can be used to visualize the results in GWASDropsxE, with a QQ-plot, Manhattan plot or QTL-plot. More details for each `plotType` are available in the statgenGWAS vignette.


#### QQ plots

```{r qqMtgxE} 
plot(GWASDropsxE, plotType = "qq")
```

#### Manhattan plots

```{r manhattanMtgxE}
plot(GWASDropsxE, plotType = "manhattan")
```

#### QTL plots

The trait is measured with the same scale across trials so the effect estimates can be compared directly (can set `normalize = FALSE`).

```{r qtlMtgNormxE}
## Set significance threshold to 6 and do not normalize effect estimates.
plot(GWASDropsxE, plotType = "qtl", yThr = 6, normalize = FALSE)
```

----

# Further options {#opts}

The *runMultiTraitGwas* function has many more parameters that can be specified. In this section similar parameters are grouped and explained with examples on how to use them.


## Significance thresholds {#sign}

The threshold for selecting significant SNPs in a GWAS analysis is computed by default using Bonferroni correction, with an alpha of 0.05. The alpha can be modified setting the option *alpha* when calling *runMultiTraitGwas*. Two other threshold types can be used: a fixed threshold (`thrType = "fixed"`) specifying the $-\log_{10}(p)$ (*LODThr*) value of the threshold, or a threshold that defines the n SNPs with the highest $-\log_{10}(p)$ scores as significant SNPs. Set `thrType = "small"` together with `nSnpLOD = n` to do this. In the following example, we select all SNPs with $p < 10^{-4}$.

```{r mtgSNPFixThr, eval = FALSE}
## Run multi-trait GWAS for Mur13W.
## Use a fixed significance threshold of 4.
GWASDropsFixThr <- runMultiTraitGwas(gData = gDataDropsDedup,
                                     trials = "Mur13W", 
                                     covModel = "fa")
```

## Variance components

There are three ways to compute the variance components used in the GWAS analysis. These can be specified by setting the parameter `covModel`. See [**2.2.2**](#vc) for a description of the options.

**Note that `covModel = unst` can only be used for less than 10 traits or trials. It is not recommended to use it for six to nine trials for computational reasons.**

```{r mtgSNPNR, eval = FALSE}
## Run multi-trait GWAS for for Mur13W.
## Use a factor analytic model for computing the variance components.
GWASDropsFA <- runMultiTraitGwas(gData = gDataDropsDedup,
                                 trials = "Mur13W",
                                 covModel = "fa")

## Rerun the analysis, using the variance components computed in the 
## previous model as inputs.
GWASDropsFA2 <- runMultiTraitGwas(gData = gDataDropsDedup,
                                  trials = "Mur13W",
                                  fitVarComp  = FALSE,
                                  Vg = GWASDropsFA$GWASInfo$varComp$Vg,
                                  Ve = GWASDropsFA$GWASInfo$varComp$Ve)
```

## Covariates

Covariates can be included as extra fixed effects in the GWAS model. The *runMultiTraitGwas* function distinguishes between 'usual' covariates and SNP-covariates. The former could be design factors such as block, or other traits one wants to condition on. In the latter case, the covariate(s) are one or more of the markers contained in the genotypic data. SNP-covariates can be set with the parameter `snpCov`, which should be a vector of marker-names. Similarly, other covariates should be specified using the parameter `covar`, containing a vector of covariate names. The gData object should contain these covariates in gData$covar.

In case SNP-covariates are used, GWAS for all the other SNPs is performed with the the SNP-covariates as extra fixed effect; also the null model used to estimate the variance components includes these effects. For each SNP in SNP-covariates, a p-value is obtained using the same F-test and null model to estimate the variance components, but with only all other SNPs (if any) in SNP-covariates as fixed effects.

```{r mtgSNPCovar, eval = FALSE}
## Run multi-trait GWAS for Mur13W.
## Use PZE-106021410, the most significant SNP, a SNP covariate.
GWASDropsSnpCov <- runMultiTraitGwas(gData = gDataDropsDedup,
                                     trials = "Mur13W",
                                     snpCov = "PZE-106021410",
                                     covModel = "fa")
```

## Minor Allele Frequency

It is recommended to remove SNPs with a low minor allele frequency (MAF) from the data before starting a GWAS analysis. However it is also possible to do so in the analysis itself. The difference between these approaches is that *codeMarkers* removes the SNPs, whereas *runMultiTraitGwas* excludes them from the analysis but leaves them in the output (with results set to NA). In the latter case it will still be possible to see the allele frequency of the SNP.  
By default all SNPs with a MAF lower than 0.01 are excluded from the analysis. This can be controlled by the parameter `MAF`. Setting MAF to 0 will still exclude duplicate SNPs since duplicates cause problems when fitting the underlying models. 

```{r mtgMAF, eval = FALSE}
## Run multi-trait GWAS for Mur13W.
## Only include SNPs that have a MAF of 0.05 or higher.
GWASDropsMAF <- runMultiTraitGwas(gData = gDataDropsDedup,
                                  trials = "Mur13W",
                                  covModel = "fa",
                                  MAF = 0.05)
```

## Estimation of common SNP effects and QTL×E effects.

Besides a normal SNP-effect model, it is possible to fit a common SNP-effect model as well (see [**2.2.1**](#common)). When doing so, in addition to the SNP-effect, also the common SNP-effect and the QTL×E effect and corresponding standard errors and P-values are returned. These are included as extra columns in the GWAResult data.table in the function output.
```{r mtgCommon, eval = FALSE}
## Run multi-trait GWAS for Mur13W.
## Fit an additional common sNP effect model.
GWASDropsCommon <- runMultiTraitGwas(gData = gDataDropsDedup,
                                     trials = "Mur13W",
                                     covModel = "fa",
                                     estCom = TRUE)
head(GWASDropsCommon$GWAResult$Mur13W)

```

----

## References
